Stochastic Approximation Monte Carlo with a Dynamic Update Factor (SAD)

### 2019-02-11 papers/histogram/sad.tex

Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for SAD, WL,
and SAMC.  The update factor
for WL remains high for many iterations, and then decreases very rapidly,
while the SAMC $\gamma_t$ remains constant before dropping as $1/t$.
The update factor for SAD in contrast fluctuates
dynamically around a value less than 1 for early MC moves.
{\color{red} For intermediate MC moves, $\gamma_t$ decreases as
$1/t$ while continuing to fluctuate whenever a new important energy
is discovered.  Once all the important energy states are discovered, the
update factor decreases faster than $1/t$ until asymptoting to $N_S/t$.
}

@@

%\begin{figure}
  %\includegraphics[width=\columnwidth]{figs/gamma-n500}
  %\caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
    %different methods: WL, SAMC, and SAD.
    %method.}\label{fig:gamma-vs-t}
%\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/gamma-n256}
  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for three
    different methods: WL, SAMC, and SAD.
    method.}\label{fig:gamma-vs-t}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/gamma-n50}
  \caption{The update factor $\gamma_t$ for $N=50$ versus iteration for three
    different methods: WL, SAMC, and SAD.}
\end{figure}

>>>>

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/gamma-n256}
  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for
    four different methods: WL, $1/t$-WL, SAMC, and SAD}
    \label{fig:gamma-vs-t}
\end{figure}
Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related methods
SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains constant
before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for many iterations,
and then decreases very rapidly, with $1/t$-WL behaving similarly before
transitioning to a $1/t$ behavior.
The update factor for SAD fluctuates
dynamically around a value less than 1 for early MC moves, and then
decreases as approximately $1/t$ while continuing to fluctuate as 
new energies are found to be important.  At late times, the SAD $\gamma_t$
decreases as approximately $1/t^2$ until asymptoting to $N_S/t$, which
is the same as $1/t$-WL.

@@

%\begin{figure}
  %\includegraphics[width=\columnwidth]{figs/gamma-n500}
  %\caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
    %different methods: WL, SAMC, and SAD.
    %method.}\label{fig:gamma-vs-t}
%\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/gamma-n50}
  \caption{The update factor $\gamma_t$ for $N=50$ versus iteration for three
    different methods: WL, SAMC, and SAD.}
\end{figure}
